Title

### 2018-09-05 thesis/paper/paper.tex
Continued evaluation

\paragraph{Rules}
The property scope constraint requires no context information to be violated; the triple, qualifier, or reference with the constrained property is enough. This means violations can be found using these rules:
\begin{align*}
\text{violation\_triple}(S, I, \text{propertyConstant}, V) &\leftarrow \text{statement\_EDB}(S, I, \text{propertyConstant}, V)\\
\text{violation\_qualifier}(S, \text{propertyConstant}, V) &\leftarrow \text{qualifier\_EDB}(S, \text{propertyConstant}, V)\\
\text{violation\_reference}(S, \text{propertyConstant}, V) &\leftarrow \text{reference\_EDB}(S, H, \text{propertyConstant}, V)
\end{align*}
Which of these rules need to be used depends on the values of the \emph{property scope} qualifier. If a concrete constrained on a property does not have \emph{as main value} as a value of the qualifier, the first rule is used. The case for \emph{as qualifier} and \emph{as reference} is analogue with the second and third rule, respectively.

@@

\paragraph{Rules}
The conflicts-with constraint is violated if another statement on the same item either has a conflicting property or a conflicting property and a conflicting value depending on specified qualifiers. It depends on a conjunction (\(O\) denoting another statement ID, \(C\) another value) and a violation for a conflicts-with constraint is thus either
\begin{equation*}
\begin{split}
\text{violation\_triple}&(S, I, \text{propertyConstant}, V) \leftarrow\\
&\text{statement\_EDB}(S, I, \text{propertyConstant}, V) \wedge{}\\
&\text{statement\_EDB}(O, I, \text{conflictingPropertyConstant}, C)
\end{split}
\end{equation*}
for constraints with no conflicting values specified or one
\begin{equation*}
\begin{split}
\text{violation\_triple}&(S, I, \text{propertyConstant}, V) \leftarrow\\
&\text{statement\_EDB}(S, I, \text{propertyConstant}, V) \wedge{}\\
&\text{statement\_EDB}(O, I, \text{conflictingPropertyConstant}, \text{conflictingValueConstant})
\end{split}
\end{equation*}
for each conflicting value if specified.\\

@@

\paragraph{Rules}
The allowed entity types constraint is violated by a statement with the wrong entity type.  Constraints regarding the lexeme entity type can be ignored since they are currently not used. The Violations depend on a simple conjunction:
\begin{equation*}
\begin{split}
\text{violation\_triple}&(S, I, \text{propertyConstant}, V) \leftarrow \\
&\text{statement\_EDB}(S, I, \text{propertyConstant}, V) \wedge \text{item}(I)
\end{split}
\end{equation*}
\begin{equation*}
\begin{split}
\text{violation\_triple}&(S, I, \text{propertyConstant}, V) \leftarrow \\
&\text{statement\_EDB}(S, I, \text{propertyConstant}, V) \wedge \text{property}(I)
\end{split}
\end{equation*}
The first rule are created for all \emph{allowed entity type} constraints whose \emph{allowed entity type} does not contain \emph{Wikibase item}. The second rule works analogue for \emph{Wikibase property}.

@@

\paragraph{Rules}
The none of constraint simply forbids certain values for properties, thus effectively prohibiting certain triples, qualifiers and references. The rules, depending on nothing but specific database entries, are
\begin{equation*}
\begin{split}
\text{violation\_triple}&(S, I, \text{propertyConstant}, \text{forbiddenValueConstant}) \leftarrow \\
&\text{statement\_EDB}(S, I, \text{propertyConstant}, \text{forbiddenValueConstant})
\end{split}
\end{equation*}
\begin{equation*}
\begin{split}
\text{violation\_qualifier}&(S, \text{propertyConstant}, \text{forbiddenValueConstant}) \leftarrow \\
&\text{qualifier\_EDB}(S, \text{propertyConstant}, \text{forbiddenValueConstant})
\end{split}
\end{equation*}
\begin{equation*}
\begin{split}
\text{violation\_reference}&(S, \text{propertyConstant}, \text{forbiddenValueConstant}) \leftarrow \\
&\text{reference\_EDB}(S, H, \text{propertyConstant}, \text{forbiddenValueConstant})
\end{split}
\end{equation*}
Given a specified constraint, each of the three rules is created once for each \emph{forbidden value}.

@@

\paragraph{Rules}
The \emph{distinct values} constraint requires two distinct constrained statements with the same value. 
Two constrained statements with the same value can be found via a simple conjunction of two statement\_EDB atoms with the propertyConstant as predicate and a variable \(V\) as value. Since this would mark every constrained statement as a violations because it can fulfil both atoms at the same time, inequality of the statement IDs \(S\) and \(O\) must be tested as well.
\begin{equation*}
\begin{split}
\text{violation\_triple}&(S, I, \text{propertyConstant}, V) \leftarrow \\
&\text{statement\_EDB}(S, I, \text{propertyConstant}, V) \wedge{} \\
&\text{statement\_EDB}(O, X, \text{propertyConstant}, V) \wedge{} \\
&\text{unequal}(S, O)
\end{split}
\end{equation*}
The rules for qualifiers and reference are analogue.

@@

Violation are thus
\begin{equation*}
\begin{split}
\text{violation\_triple}&(S, I, \text{propertyConstant}, V) \leftarrow \\
&\text{statement\_EDB}(S, I, \text{propertyConstant}, V) \wedge{} \\
&\text{unit}(V, U) \wedge \text{unequal}(\{A\}, U)
\end{split}
\end{equation*}
\begin{equation*}
\begin{split}
\text{violation\_qualifier}&(S, I, \text{propertyConstant}, V) \leftarrow \\
&\text{qualifier\_EDB}(S, \text{propertyConstant}, V) \wedge{} \\
&\text{unit}(V, U) \wedge \text{unequal}(\{A\}, U)
\end{split}
\end{equation*}
\begin{equation*}
\begin{split}
\text{violation\_reference}&(S, I, \text{propertyConstant}, V) \leftarrow \\
&\text{reference\_EDB}(S, H, \text{propertyConstant}, V) \wedge{} \\
&\text{unit}(V, U) \wedge \text{unequal}(\{A\}, U)
\end{split}
\end{equation*}

@@

\paragraph{Rules}
To represent a violation of the \emph{allowed qualifiers} constraint in datalog inequality between all specified allowed qualifiers and the properties of qualifiers is required. The constraint is violated if a qualifier on a constrained statement exists that is unequal to all allowed qualifiers. The notation for set inequality as introduced in section \ref{subsec_3_allowed_units} will be used. With \(A\) the set of allowed qualifiers the rule is:
\begin{equation*}
\begin{split}
\text{violation\_triple}&(S, I, \text{propertyConstant}, V) \leftarrow \\
&\text{statement\_EDB}(S, I, \text{propertyConstant}, V) \wedge{} \\
&\text{qualifier\_EDB}(S, Q, O) \wedge \text{unequal}(\{A\}, Q)
\end{split}
\end{equation*}

@@

\paragraph{Rules}
The one-of constraint fits in the pattern established by the previous two constraints and disallows certain property-value combinations in statements, qualifiers and references. Inequality is established between all \emph{allowed values} and the values of statements, qualifiers and references, respectively. With \(A\) the set of allowed values the rules for triple, qualifier and reference are:
\begin{align*}
\text{violation\_triple}(S, I, \text{propertyConstant}, V) &\leftarrow \\
\text{statement\_EDB}(S, I, &\text{propertyConstant}, V) \wedge \text{unequal}(\{A\}, V) \\
\text{violation\_qualifier}(S, \text{propertyConstant}, V) &\leftarrow \\
\text{qualifier\_EDB}(S, &\text{propertyConstant}, V) \wedge \text{unequal}(\{A\}, V) \\
\text{violation\_reference}(S, \text{propertyConstant}, V) &\leftarrow \\
\text{reference\_EDB}(S, H, &\text{propertyConstant}, V) \wedge \text{unequal}(\{A\}, V)
\end{align*}

@@

\paragraph{Rules}
The \emph{item requires statement} constraint requires inequality between all properties of statements and the \emph{required property} as well as between all statement values and the \emph{allowed values}. The order of the item statements needs to be established. An item violates the constraint if it a) has a statement using constrained property as predicate, b1) has no statement using the \emph{required property} as predicate, or (if \emph{allowed values} are specified) b2) has only statement using the \emph{required property} as predicate that do not use one of the \emph{allowed values} as value. The required term \(t\) is the \emph{required property}. Without separators the conjunction is
\begin{equation*}
\begin{split}
c_1 = &\text{statement\_EDB}(Q, I, \text{propertyConstant}, X) \wedge{} \\
&\text{statement\_EDB}(S, I, P, V) \wedge \text{unequal}(\text{requiredPropertyConstant}, P)
\end{split}
\end{equation*} \(\)
If a set of \emph{allowed values} \(A\) has been specified, a second set of require-rules have to be introduced with
\begin{equation*}
\begin{split}
c_2 = &\text{statement\_EDB}(Q, I, \text{propertyConstant}, X) \wedge{} \\
&\text{statement\_EDB}(S, I, \text{requiredPropertyConstant}, V) \wedge \text{unequal}(\{A\}, V)
\end{split}
\end{equation*}
Note that the first set of rules is still required, since the preceding statements do not need to have the \emph{required property}. In both cases violations can be found with
\begin{equation*}
\begin{split}
\text{violation\_triple}&(S, I, \text{propertyConstant}, V) \leftarrow \\
&\text{statement\_EDB}(S, I, \text{propertyConstant}, V) \wedge{} \\
&\text{statement\_EDB}(O, I, P, X) \wedge \text{last}(O, I) \wedge{} \\
&\text{require}(O, \text{propertyConstant}, \text{requiredPropertyConstant})
\end{split}
\end{equation*}

@@

\paragraph{Rules}
The \emph{value requires statement} constraint is syntactically very similar to the previous \emph{item requires statement} constraint. Instead of requiring a statement on the item of a constrained statement it requires it on the value. Inequality need to be computed between all properties and all values and the item statement order established. The conjunctions change in the first statement and are:
\begin{equation*}
\begin{split}
c_1 = &\text{statement\_EDB}(Q, R, \text{propertyConstant}, I) \wedge{} \\
&\text{statement\_EDB}(S, I, P, V) \wedge \text{unequal}(\text{requiredPropertyConstant}, P)
\end{split}
\end{equation*} \(\)
\begin{equation*}
\begin{split}
c_2 = &\text{statement\_EDB}(Q, R, \text{propertyConstant}, I) \wedge{} \\
&\text{statement\_EDB}(S, I, \text{requiredPropertyConstant}, V) \wedge \text{unequal}(\{A\}, V)
\end{split}
\end{equation*}
The final violation rule would then be:
\begin{equation*}
\begin{split}
\text{violation\_triple}&(S, I, \text{propertyConstant}, V) \leftarrow \\
&\text{statement\_EDB}(S, I, \text{propertyConstant}, V) \wedge{} \\
&\text{statement\_EDB}(O, V, P, X) \wedge \text{last}(O, V) \wedge{} \\
&\text{require}(O, \text{propertyConstant}, \text{requiredPropertyConstant})
\end{split}
\end{equation*}

@@

\subsubsection{No separators}
If the separator set is empty, the violation simply occurs if there are two constrained statements on the same item. 
\begin{equation}\label{eq_no_separators_triple}
\begin{split}
\text{violation\_triple}&(S, I, \text{propertyConstant}, V) \leftarrow \\
&\text{statement\_EDB}(S, I, \text{propertyConstant}, V) \wedge{} \\
&\text{statement\_EDB}(O, I, \text{propertyConstant}, X) \wedge{} \\
&\text{unequal}(S, O)
\end{split}
\end{equation}
For qualifiers and references there is no ID whose inequality could be checked. However, since the statement ID and the property of a possible conflict are always the same, inequality can be tested on the values.
\begin{equation*}
\begin{split}
\text{violation\_qualifier}&(S, \text{propertyConstant}, V) \leftarrow \\
&\text{qualifier\_EDB}(S, \text{propertyConstant}, V) \wedge{} \\
&\text{qualifier\_EDB}(S, \text{propertyConstant}, X) \wedge{} \\
&\text{unequal}(V, X)
\end{split}
\end{equation*}
\begin{equation*}
\begin{split}
\text{violation\_reference}&(S, H, \text{propertyConstant}, V) \leftarrow \\
&\text{reference\_EDB}(S, H, \text{propertyConstant}, V) \wedge{} \\
&\text{reference\_EDB}(S, J, \text{propertyConstant}, X) \wedge{} \\
&\text{unequal}(V, X)
\end{split}
\end{equation*}

@@

With qualifier non-existence established the violation can be found. Two statements form a violation if they are distinct as per equation \ref{eq_no_separators_triple} and additionally fulfil the condition that all values of \emph{separator} qualifiers are equal or no such \emph{separator} qualifier exist. To determine this a number of rules have to be introduced which are based on two sets the \emph{separators} are split into: \(X\) contains the \emph{separators} that exist as qualifiers and \(Y\) the \emph{separators} that do not. The following rule is then created for every possible split of the \emph{separators} into two such sets. Note that \(\text{predicate}(\{X\})\) denotes \(\bigwedge_{x \in X} \text{predicate}(x)\) as introduced in section \ref{subsec_3_allowed_units}.
\begin{equation*}
\begin{split}
\text{violation\_triple}&(S, I, \text{propertyConstant}, V) \leftarrow \\
&\text{statement\_EDB}(S, I, \text{propertyConstant}, V) \wedge{} \\
&\text{statement\_EDB}(O, I, \text{propertyConstant}, C) \wedge{} \\
&\text{unequal}(S, O) \wedge{} \\
&\text{has\_same}(S, O, \{X\}) \wedge{} \\
&\text{does\_not\_have}(S, \{Y\}) \wedge{} \\
&\text{does\_not\_have}(O, \{Y\})
\end{split}
\end{equation*}
The predicate does\_not\_have\([2]\) linking a statement ID to the separator it does not have via 
\begin{equation*}
\begin{split}
\text{does\_not\_have}(S, R) \leftarrow &\text{statement\_EDB}(S, I, \text{propertyConstant}, C) \wedge{} \\
&\text{last\_qualifier}(S, P, V) \wedge \text{require}(S, P, V, \text{propertyConstant}, R)
\end{split}
\end{equation*}
and the predicate has\_same[2] linking two statement IDs to the separator for which the values are the same via
\begin{equation*}
\begin{split}
\text{has\_same}(S, O, Q) \leftarrow &\text{statement\_EDB}(S, I, \text{propertyConstant}, C) \wedge{} \\
&\text{statement\_EDB}(O, I, \text{propertyConstant}, X) \wedge{} \\
&\text{qualifier\_EDB}(S, Q, V) \wedge \text{qualifier\_EDB}(O, Q, V)
\end{split}
\end{equation*}
are introduced for brevity.

@@

\paragraph{Rules}
The \emph{mandatory qualifier} constraint needs to determine qualifier non-existence on constrained statements to find violations, which is similar to section \ref{subsubsec_with_separators}. The qualifier order needs to be provided and inequality must be established between all qualifier properties. The conjunction for the require-pattern is:
\begin{equation*}
\begin{split}
c = &\text{statement\_EDB}(S, I, \text{propertyConstant}, C) \wedge{} \\
&\text{qualifier\_EDB}(S, P, V) \wedge \text{unequal}(P, \text{requiredPropertyConstant})
\end{split}
\end{equation*}
With the usual require-pattern for qualifiers and the does\_not\_have rule as introduced in section \ref{subsubsec_with_separators} added for brevity the violations would be found with:
\begin{equation*}
\begin{split}
\text{violation\_triple}&(S, I, \text{propertyConstant}, V) \leftarrow \\
&\text{statement\_EDB}(S, I, \text{propertyConstant}, V) \wedge \\
&\text{does\_not\_have}(S, \text{requiredPropertyConstant})
\end{split}
\end{equation*}

@@

\subsubsection{No separators}
If the \emph{separator} set is empty, the violation simply occurs if there are two constrained statements on the same item with the preferred rank. Thus the rule \eqref{eq_no_separators_triple} is expanded to
\begin{equation*}
\begin{split}
\text{violation\_triple}&(S, I, \text{propertyConstant}, V) \leftarrow \\
&\text{statement\_EDB}(S, I, \text{propertyConstant}, V) \wedge{} \\
&\text{rank\_EDB}(S, \text{preferredRankConstant}) \wedge{} \\
&\text{statement\_EDB}(O, I, \text{propertyConstant}, X) \wedge{} \\
&\text{rank\_EDB}(O, \text{preferredRankConstant}) \wedge{} \\
&\text{unequal}(S, O)
\end{split}
\end{equation*}

\subsubsection{With separators}
As in the case above, the qualifier non-existence and violation rules are very similar to the \emph{single value} constraint. The only change is the addition of the rank requirements as shown below.
\begin{equation*}
\begin{split}
\text{violation\_triple}&(S, I, \text{propertyConstant}, V) \leftarrow \\
&\text{statement\_EDB}(S, I, \text{propertyConstant}, V) \wedge{} \\
&\text{rank\_EDB}(S, \text{preferredRankConstant}) \wedge{} \\
&\text{statement\_EDB}(O, I, \text{propertyConstant}, W) \wedge{} \\
&\text{rank\_EDB}(O, \text{preferredRankConstant}) \wedge{} \\
&\text{unequal}(S, O) \wedge{} \\
&\text{has\_same}(S, O, \{X\}) \wedge{} \\
&\text{does\_not\_have}(S, \{Y\}) \wedge{} \\
&\text{does\_not\_have}(O, \{Y\})
\end{split}
\end{equation*}

@@

{\bf Total Time} is the execution time taken for evaluating one constraint. This does not include the data extraction from JSON, nor the generation of the files necessary to establish inequality. It does include the time taken to generate the rules as well as to load, reason and query with VLog.

@@

{\bf Derivations} counts how many new facts were created during materialization. This will include all found violations but can be more in cases where additional facts were created as helpers.

@@

Most constraints could be evaluated in under two hours on even the largest dataset. The exceptions are \emph{distinct values}, \emph{item requires statement} and \emph{single value}. In parts this is expected behaviour, since these three constraints are applied to an order of magnitude more properties than all other properties with the exception of \emph{scope}, which is described in very simple rules. However, as seen in Figure \ref{fig_runtime}, the runtimes of these three constraints are multiple orders of magnitude higher than those of other constraints. The following sections will thus focus on the comparison of these three constraints to identify bottlenecks and reasons for this huge difference in performance.

>>>>

\paragraph{Rules}
The property scope constraint requires no context information to be violated; the triple, qualifier, or reference with the constrained property is enough. This means violations can be found using these rules:
\begin{align*}
\text{violation\_statement}(S, I, \text{propertyConstant}, V) &\leftarrow \text{statement\_EDB}(S, I, \text{propertyConstant}, V)\\
\text{violation\_qualifier}(S, \text{propertyConstant}, V) &\leftarrow \text{qualifier\_EDB}(S, \text{propertyConstant}, V)\\
\text{violation\_reference}(S, \text{propertyConstant}, V) &\leftarrow \text{reference\_EDB}(S, H, \text{propertyConstant}, V)
\end{align*}
Which of these rules need to be used depends on the values of the \emph{property scope} qualifier. If a concrete constrained on a property does not have \emph{as main value} as a value of the qualifier, the first rule is used. The case for \emph{as qualifier} and \emph{as reference} is analogue with the second and third rule, respectively.

@@

\paragraph{Rules}
The conflicts-with constraint is violated if another statement on the same item either has a conflicting property or a conflicting property and a conflicting value depending on specified qualifiers. It depends on a conjunction (\(O\) denoting another statement ID, \(C\) another value) and a violation for a conflicts-with constraint is thus either
\begin{equation*}
\begin{split}
\text{violation\_statement}&(S, I, \text{propertyConstant}, V) \leftarrow\\
&\text{statement\_EDB}(S, I, \text{propertyConstant}, V) \wedge{}\\
&\text{statement\_EDB}(O, I, \text{conflictingPropertyConstant}, C)
\end{split}
\end{equation*}
for constraints with no conflicting values specified or one
\begin{equation*}
\begin{split}
\text{violation\_statement}&(S, I, \text{propertyConstant}, V) \leftarrow\\
&\text{statement\_EDB}(S, I, \text{propertyConstant}, V) \wedge{}\\
&\text{statement\_EDB}(O, I, \text{conflictingPropertyConstant}, \text{conflictingValueConstant})
\end{split}
\end{equation*}
for each conflicting value if specified.\\

@@

\paragraph{Rules}
The allowed entity types constraint is violated by a statement with the wrong entity type.  Constraints regarding the lexeme entity type can be ignored since they are currently not used. The Violations depend on a simple conjunction:
\begin{equation*}
\begin{split}
\text{violation\_statement}&(S, I, \text{propertyConstant}, V) \leftarrow \\
&\text{statement\_EDB}(S, I, \text{propertyConstant}, V) \wedge \text{item}(I)
\end{split}
\end{equation*}
\begin{equation*}
\begin{split}
\text{violation\_statement}&(S, I, \text{propertyConstant}, V) \leftarrow \\
&\text{statement\_EDB}(S, I, \text{propertyConstant}, V) \wedge \text{property}(I)
\end{split}
\end{equation*}
The first rule are created for all \emph{allowed entity type} constraints whose \emph{allowed entity type} does not contain \emph{Wikibase item}. The second rule works analogue for \emph{Wikibase property}.

@@

\paragraph{Rules}
The none of constraint simply forbids certain values for properties, thus effectively prohibiting certain triples, qualifiers and references. The rules, depending on nothing but specific database entries, are
\begin{equation*}
\begin{split}
\text{violation\_statement}&(S, I, \text{propertyConstant}, \text{forbiddenValueConstant}) \leftarrow \\
&\text{statement\_EDB}(S, I, \text{propertyConstant}, \text{forbiddenValueConstant})
\end{split}
\end{equation*}
\begin{equation*}
\begin{split}
\text{violation\_qualifier}&(S, \text{propertyConstant}, \text{forbiddenValueConstant}) \leftarrow \\
&\text{qualifier\_EDB}(S, \text{propertyConstant}, \text{forbiddenValueConstant})
\end{split}
\end{equation*}
\begin{equation*}
\begin{split}
\text{violation\_reference}&(S, \text{propertyConstant}, \text{forbiddenValueConstant}) \leftarrow \\
&\text{reference\_EDB}(S, H, \text{propertyConstant}, \text{forbiddenValueConstant})
\end{split}
\end{equation*}
Given a specified constraint, each of the three rules is created once for each \emph{forbidden value}.

@@

\paragraph{Rules}
The \emph{distinct values} constraint requires two distinct constrained statements with the same value. 
Two constrained statements with the same value can be found via a simple conjunction of two statement\_EDB atoms with the propertyConstant as predicate and a variable \(V\) as value. Since this would mark every constrained statement as a violations because it can fulfil both atoms at the same time, inequality of the statement IDs \(S\) and \(O\) must be tested as well.
\begin{equation*}
\begin{split}
\text{violation\_statement}&(S, I, \text{propertyConstant}, V) \leftarrow \\
&\text{statement\_EDB}(S, I, \text{propertyConstant}, V) \wedge{} \\
&\text{statement\_EDB}(O, X, \text{propertyConstant}, V) \wedge{} \\
&\text{unequal}(S, O)
\end{split}
\end{equation*}
The rules for qualifiers and reference are analogue.

@@

Violation are thus
\begin{equation*}
\begin{split}
\text{violation\_statement}&(S, I, \text{propertyConstant}, V) \leftarrow \\
&\text{statement\_EDB}(S, I, \text{propertyConstant}, V) \wedge{} \\
&\text{unit}(V, U) \wedge \text{unequal}(\{A\}, U)
\end{split}
\end{equation*}
\begin{equation*}
\begin{split}
\text{violation\_qualifier}&(S, I, \text{propertyConstant}, V) \leftarrow \\
&\text{qualifier\_EDB}(S, \text{propertyConstant}, V) \wedge{} \\
&\text{unit}(V, U) \wedge \text{unequal}(\{A\}, U)
\end{split}
\end{equation*}
\begin{equation*}
\begin{split}
\text{violation\_reference}&(S, I, \text{propertyConstant}, V) \leftarrow \\
&\text{reference\_EDB}(S, H, \text{propertyConstant}, V) \wedge{} \\
&\text{unit}(V, U) \wedge \text{unequal}(\{A\}, U)
\end{split}
\end{equation*}

@@

\paragraph{Rules}
To represent a violation of the \emph{allowed qualifiers} constraint in datalog inequality between all specified allowed qualifiers and the properties of qualifiers is required. The constraint is violated if a qualifier on a constrained statement exists that is unequal to all allowed qualifiers. The notation for set inequality as introduced in section \ref{subsec_3_allowed_units} will be used. With \(A\) the set of allowed qualifiers the rule is:
\begin{equation*}
\begin{split}
\text{violation\_statement}&(S, I, \text{propertyConstant}, V) \leftarrow \\
&\text{statement\_EDB}(S, I, \text{propertyConstant}, V) \wedge{} \\
&\text{qualifier\_EDB}(S, Q, O) \wedge \text{unequal}(\{A\}, Q)
\end{split}
\end{equation*}

@@

\paragraph{Rules}
The one-of constraint fits in the pattern established by the previous two constraints and disallows certain property-value combinations in statements, qualifiers and references. Inequality is established between all \emph{allowed values} and the values of statements, qualifiers and references, respectively. With \(A\) the set of allowed values the rules for triple, qualifier and reference are:
\begin{align*}
\text{violation\_statement}(S, I, \text{propertyConstant}, V) &\leftarrow \\
\text{statement\_EDB}(S, I, &\text{propertyConstant}, V) \wedge \text{unequal}(\{A\}, V) \\
\text{violation\_qualifier}(S, \text{propertyConstant}, V) &\leftarrow \\
\text{qualifier\_EDB}(S, &\text{propertyConstant}, V) \wedge \text{unequal}(\{A\}, V) \\
\text{violation\_reference}(S, \text{propertyConstant}, V) &\leftarrow \\
\text{reference\_EDB}(S, H, &\text{propertyConstant}, V) \wedge \text{unequal}(\{A\}, V)
\end{align*}

@@

\paragraph{Rules}
The \emph{item requires statement} constraint requires inequality between all properties of statements and the \emph{required property} as well as between all statement values and the \emph{allowed values}. The order of the item statements needs to be established. An item violates the constraint if it a) has a statement using constrained property as predicate, b1) has no statement using the \emph{required property} as predicate, or (if \emph{allowed values} are specified) b2) has only statement using the \emph{required property} as predicate that do not use one of the \emph{allowed values} as value. The required term \(t\) is the \emph{required property}. Without separators the conjunction is
\begin{equation*}
\begin{split}
c_1 = &\text{statement\_EDB}(Q, I, \text{propertyConstant}, X) \wedge{} \\
&\text{statement\_EDB}(S, I, P, V) \wedge \text{unequal}(\text{requiredPropertyConstant}, P)
\end{split}
\end{equation*} \(\)
If a set of \emph{allowed values} \(A\) has been specified, a second set of require-rules have to be introduced with
\begin{equation*}
\begin{split}
c_2 = &\text{statement\_EDB}(Q, I, \text{propertyConstant}, X) \wedge{} \\
&\text{statement\_EDB}(S, I, \text{requiredPropertyConstant}, V) \wedge \text{unequal}(\{A\}, V)
\end{split}
\end{equation*}
Note that the first set of rules is still required, since the preceding statements do not need to have the \emph{required property}. In both cases violations can be found with
\begin{equation*}
\begin{split}
\text{violation\_statement}&(S, I, \text{propertyConstant}, V) \leftarrow \\
&\text{statement\_EDB}(S, I, \text{propertyConstant}, V) \wedge{} \\
&\text{statement\_EDB}(O, I, P, X) \wedge \text{last}(O, I) \wedge{} \\
&\text{require}(O, \text{propertyConstant}, \text{requiredPropertyConstant})
\end{split}
\end{equation*}

@@

\paragraph{Rules}
The \emph{value requires statement} constraint is syntactically very similar to the previous \emph{item requires statement} constraint. Instead of requiring a statement on the item of a constrained statement it requires it on the value. Inequality need to be computed between all properties and all values and the item statement order established. The conjunctions change in the first statement and are:
\begin{equation*}
\begin{split}
c_1 = &\text{statement\_EDB}(Q, R, \text{propertyConstant}, I) \wedge{} \\
&\text{statement\_EDB}(S, I, P, V) \wedge \text{unequal}(\text{requiredPropertyConstant}, P)
\end{split}
\end{equation*} \(\)
\begin{equation*}
\begin{split}
c_2 = &\text{statement\_EDB}(Q, R, \text{propertyConstant}, I) \wedge{} \\
&\text{statement\_EDB}(S, I, \text{requiredPropertyConstant}, V) \wedge \text{unequal}(\{A\}, V)
\end{split}
\end{equation*}
The final violation rule would then be:
\begin{equation*}
\begin{split}
\text{violation\_statement}&(S, I, \text{propertyConstant}, V) \leftarrow \\
&\text{statement\_EDB}(S, I, \text{propertyConstant}, V) \wedge{} \\
&\text{statement\_EDB}(O, V, P, X) \wedge \text{last}(O, V) \wedge{} \\
&\text{require}(O, \text{propertyConstant}, \text{requiredPropertyConstant})
\end{split}
\end{equation*}

@@

\subsubsection{No separators}
If the separator set is empty, the violation simply occurs if there are two constrained statements on the same item. 
\begin{equation}\label{eq_no_separators_triple}
\begin{split}
\text{violation\_statement}&(S, I, \text{propertyConstant}, V) \leftarrow \\
&\text{statement\_EDB}(S, I, \text{propertyConstant}, V) \wedge{} \\
&\text{statement\_EDB}(O, I, \text{propertyConstant}, X) \wedge{} \\
&\text{unequal}(S, O)
\end{split}
\end{equation}
For qualifiers and references there is no ID whose inequality could be checked. However, since the statement ID and the property of a possible conflict are always the same, inequality can be tested on the values.
\begin{equation*}
\begin{split}
\text{violation\_qualifier}&(S, \text{propertyConstant}, V) \leftarrow \\
&\text{qualifier\_EDB}(S, \text{propertyConstant}, V) \wedge{} \\
&\text{qualifier\_EDB}(S, \text{propertyConstant}, X) \wedge{} \\
&\text{unequal}(V, X)
\end{split}
\end{equation*}
\begin{equation*}
\begin{split}
\text{violation\_reference}&(S, H, \text{propertyConstant}, V) \leftarrow \\
&\text{reference\_EDB}(S, H, \text{propertyConstant}, V) \wedge{} \\
&\text{reference\_EDB}(S, J, \text{propertyConstant}, X) \wedge{} \\
&\text{unequal}(V, X)
\end{split}
\end{equation*}

@@

With qualifier non-existence established the violation can be found. Two statements form a violation if they are distinct as per equation \ref{eq_no_separators_triple} and additionally fulfil the condition that all values of \emph{separator} qualifiers are equal or no such \emph{separator} qualifier exist. To determine this a number of rules have to be introduced which are based on two sets the \emph{separators} are split into: \(X\) contains the \emph{separators} that exist as qualifiers and \(Y\) the \emph{separators} that do not. The following rule is then created for every possible split of the \emph{separators} into two such sets. Note that \(\text{predicate}(\{X\})\) denotes \(\bigwedge_{x \in X} \text{predicate}(x)\) as introduced in section \ref{subsec_3_allowed_units}.
\begin{equation*}
\begin{split}
\text{violation\_statement}&(S, I, \text{propertyConstant}, V) \leftarrow \\
&\text{statement\_EDB}(S, I, \text{propertyConstant}, V) \wedge{} \\
&\text{statement\_EDB}(O, I, \text{propertyConstant}, C) \wedge{} \\
&\text{unequal}(S, O) \wedge{} \\
&\text{has\_same}(S, O, \{X\}) \wedge{} \\
&\text{does\_not\_have}(S, \{Y\}) \wedge{} \\
&\text{does\_not\_have}(O, \{Y\})
\end{split}
\end{equation*}
The predicate does\_not\_have\([2]\) linking a statement ID to the separator it does not have via 
\begin{equation*}
\begin{split}
\text{does\_not\_have}(S, R) \leftarrow &\text{statement\_EDB}(S, I, \text{propertyConstant}, C) \wedge{} \\
&\text{last\_qualifier}(S, P, V) \wedge \text{require}(S, P, V, \text{propertyConstant}, R)
\end{split}
\end{equation*}
and the predicate has\_same[2] linking two statement IDs to the separator for which the values are the same via
\begin{equation*}
\begin{split}
\text{has\_same}(S, O, Q) \leftarrow &\text{statement\_EDB}(S, I, \text{propertyConstant}, C) \wedge{} \\
&\text{statement\_EDB}(O, I, \text{propertyConstant}, X) \wedge{} \\
&\text{qualifier\_EDB}(S, Q, V) \wedge \text{qualifier\_EDB}(O, Q, V)
\end{split}
\end{equation*}
are introduced for brevity.

@@

\paragraph{Rules}
The \emph{mandatory qualifier} constraint needs to determine qualifier non-existence on constrained statements to find violations, which is similar to section \ref{subsubsec_with_separators}. The qualifier order needs to be provided and inequality must be established between all qualifier properties. The conjunction for the require-pattern is:
\begin{equation*}
\begin{split}
c = &\text{statement\_EDB}(S, I, \text{propertyConstant}, C) \wedge{} \\
&\text{qualifier\_EDB}(S, P, V) \wedge \text{unequal}(P, \text{requiredPropertyConstant})
\end{split}
\end{equation*}
With the usual require-pattern for qualifiers and the does\_not\_have rule as introduced in section \ref{subsubsec_with_separators} added for brevity the violations would be found with:
\begin{equation*}
\begin{split}
\text{violation\_statement}&(S, I, \text{propertyConstant}, V) \leftarrow \\
&\text{statement\_EDB}(S, I, \text{propertyConstant}, V) \wedge \\
&\text{does\_not\_have}(S, \text{requiredPropertyConstant})
\end{split}
\end{equation*}

@@

\subsubsection{No separators}
If the \emph{separator} set is empty, the violation simply occurs if there are two constrained statements on the same item with the preferred rank. Thus the rule \eqref{eq_no_separators_triple} is expanded to
\begin{equation*}
\begin{split}
\text{violation\_statement}&(S, I, \text{propertyConstant}, V) \leftarrow \\
&\text{statement\_EDB}(S, I, \text{propertyConstant}, V) \wedge{} \\
&\text{rank\_EDB}(S, \text{preferredRankConstant}) \wedge{} \\
&\text{statement\_EDB}(O, I, \text{propertyConstant}, X) \wedge{} \\
&\text{rank\_EDB}(O, \text{preferredRankConstant}) \wedge{} \\
&\text{unequal}(S, O)
\end{split}
\end{equation*}

\subsubsection{With separators}
As in the case above, the qualifier non-existence and violation rules are very similar to the \emph{single value} constraint. The only change is the addition of the rank requirements as shown below.
\begin{equation*}
\begin{split}
\text{violation\_statement}&(S, I, \text{propertyConstant}, V) \leftarrow \\
&\text{statement\_EDB}(S, I, \text{propertyConstant}, V) \wedge{} \\
&\text{rank\_EDB}(S, \text{preferredRankConstant}) \wedge{} \\
&\text{statement\_EDB}(O, I, \text{propertyConstant}, W) \wedge{} \\
&\text{rank\_EDB}(O, \text{preferredRankConstant}) \wedge{} \\
&\text{unequal}(S, O) \wedge{} \\
&\text{has\_same}(S, O, \{X\}) \wedge{} \\
&\text{does\_not\_have}(S, \{Y\}) \wedge{} \\
&\text{does\_not\_have}(O, \{Y\})
\end{split}
\end{equation*}

@@

@@

{\bf Additional Derivations} counts how many new facts were created during materialization, excluding the violations themselves.

@@

Because of a bug in VLog\footnote{\url{https://github.com/knowsys/vlog4j/issues/43}}, the largest dataset that could be completely evaluated contained 100,000 entities. With datasets of larger size, more than half of the constraints were affected by this bug, meaning the analysis could not be conducted. All further statistics describe the results on the 100,000 entity set unless specifically noted otherwise. Since inequality could only be computed in demanded mode (see section  \ref{sec_inequality_evaluation}), all statistics will show the results attained in this mode.
